Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-23 Draft.Rmd
edited
added rf code

```{r}
library(tidyverse)
library(tidymodels)
library(broom)
library(xgboost)
library(caret)
library(plotROC)
library(randomForest)
library(factoextra)
library(MASS)

@@

## Predictor/Feature Selection
```{r}
# Use Random Forest for variable selection
rfmodel = randomForest(value~., data=train[,2:50])
import=as.data.frame(importance(rfmodel)) |> arrange(desc(IncNodePurity))
import=tibble::rownames_to_column(import, 'variables')

@@

## Models 
```{r Linear}
set.seed(123)
# Fit model and get summary
fit_lm = lm(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data=train)
step <- stepAIC(fit_lm, direction="backward", trace=TRUE)
# Drops county_area, popdens_county, and log_nei_2008_pm25_sum_10000

@@

```{r kNN ALL MODELS FAILED}
set.seed(123)
# Find ideal k neighbors
sqrt(nrow(train))

@@

```{r Xgboost}
set.seed(123)
# Preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = training(dat_split)) %>%
  # Convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # Combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # Remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

@@

xgboost_wf <- workflows::workflow() %>%
  add_model(xgboost_model) %>% 
  add_formula(value~lon+lat)

# Hyperparameter tuning
# Takes a long time to run, 5 min
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq),
  control = tune::control_grid(verbose = TRUE))

@@

# Eval. model performance on test data
train_processed <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction <- xgboost_model_final %>%
  # Fit the model on all the training data
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = train_processed) %>%
  # Predict the sale prices for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(dat_split))

@@

```{r Random Forest}
set.seed(123)
# Create Random Forest Model
# Use grid of tuning parameters
model4 <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")
# Create workflow
  wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model4)

# Fit the model on the full training dataset using the `fit()` function
res4 <- fit(wf4, data = train_scaled)
# Check performance on the complete training data
res4 %>% 
    extract_fit_engine() %>% 
    summary()
# Check performance using cross-validation
folds4 <- vfold_cv(train, v = 10)
folds4
# Determine performance of model by resampling
res4 <- fit_resamples(wf4, folds4) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res4 %>% 
    collect_metrics()

## Try a grid of tuning parameters
model <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")

wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model)

## Fit model over grid of tuning parameters
res <- tune_grid(wf4, resamples = folds4, grid = 20)

res %>% 
    collect_metrics()
```

>>>>

```{r}
library(tidyverse)
library(tidymodels)
library(broom)
library(xgboost)
library(caret)
library(plotROC)
library(randomForest)
library(factoextra)
library(MASS)
library(ranger)

@@

## Predictor/Feature Selection
```{r}
# Use Random Forest for variable selection
rfmodel = randomForest(value~., data=train[,2:50])
import=as.data.frame(randomForest::importance(rfmodel)) |> arrange(desc(IncNodePurity))
import=tibble::rownames_to_column(import, 'variables')

@@

## Models 
```{r LinearRegression}
set.seed(123)
# Fit model and get summary
fit_lm = lm(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data=train)
step <- stepAIC(fit_lm, direction="backward", trace=TRUE)
# Drops county_area, popdens_county, and log_nei_2008_pm25_sum_10000

@@

```{r kNearestNeighbors}
# kNN uses scaled data called train_knn
set.seed(123)
# Find ideal k neighbors
sqrt(nrow(train))

@@

```{r Xgboost}
set.seed(123)
# Preprocessing recipe
preprocessing_recipe <- 
  recipes::recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = training(dat_split)) %>%
  # Convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # Combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # Remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

@@

xgboost_wf <- workflows::workflow() %>%
  add_model(xgboost_model) %>% 
  add_formula(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000)

# Hyperparameter tuning
# Takes a long time to run, 5-8 min
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq),
  control = tune::control_grid(verbose = TRUE))

@@

# Eval. model performance on train data
train_processed <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction <- xgboost_model_final %>%
  # Fit the model on all the training data
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = train_processed) %>%
  # Predict value for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(dat_split))

@@

```{r RandomForest}
set.seed(123)
# Preprocessing recipe
rec4 <- recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data=train) %>%
  step_log(all_outcomes()) |> 
  step_normalize(all_numeric_predictors()) |> 
  step_dummy(all_nominal_predictors())

#  Create model specification and wf
rf_spec <- rand_forest(mtry = tune(), trees = tune(), min_n = tune()) %>%
  set_engine("ranger", num.threads=7, importance="impurity") |> 
  set_mode("regression")

wf4 <- workflow() %>%
  add_recipe(rec4) %>%
  add_model(rf_spec)

folds4 = recipes::bake(preprocessing_recipe, new_data = training(dat_split)) %>%  
  rsample::vfold_cv(v = 5)

# Space-filling designs grid
rf_grid <- grid_latin_hypercube(min_n(), mtry(range = c(4, 9)), trees(), size = 60)

# Tuning hyperparameters
library(doParallel)
all_cores = parallel::detectCores(logical=FALSE)
cl = makePSOCKcluster(all_cores)
doParallel::registerDoParallel(cl)

# Takes long time to load 2 min
tune_res <- wf4 |> tune_grid(resamples = folds4, grid = rf_grid, metrics=metric_set(rmse,rsq), control = tune::control_grid(verbose = TRUE))
tune_res

# Find hyper parameter values which performed best at minimizing RMSE
tune_res %>% tune::show_best(metric = "rmse") %>%knitr::kable()
# Isolate the best performing hyperparameter values.
rf_best_params <- tune_res %>% tune::select_best("rmse")
rf_best_params

# Final Random Forest model
rf_model_final <- rf_spec %>% finalize_model(rf_best_params)

# Check Random Forest model performance
train_processed2 <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction2 = rf_model_final |> 
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data=train_processed2) |> 
  predict(new_data=train_processed2) |> 
  bind_cols(training(dat_split))

rf_score_train = train_prediction2 |> 
  yardstick::metrics(value, .pred) |> 
  mutate(.estimate=format(round(.estimate, 2), big.mark=","))
rf_score_train
```
